Carryover effects in free recall reveal how prior experiences influence memories of new experiences

### 2023-01-06 paper/main.tex
draft of order manipulation conditions:   - added Allison as an author (yay!)   - added "stick" figure to illustrate stochastic word ordering procedure   - added descriptions + equations for order manipulation conditions

\author{Jeremy R. Manning\textsuperscript{1, *}, Kirsten
Ziman\textsuperscript{1, 2}, Emily Whitaker\textsuperscript{1},\\Paxton C.
Fitzpatrick\textsuperscript{1}, Madeline R. Lee\textsuperscript{1}, Bryan J.
Bollinger\textsuperscript{1},\\and Andrew C. Heusser\textsuperscript{1,
3}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Princeton
University\\\textsuperscript{3}Akili
Interactive\\\textsuperscript{*}Corresponding author:
jeremy.r.manning@dartmouth.edu}

@@

Each of six \textit{order manipulation} conditions used a different
feature-based sorting procedure to order words on early lists, where each
sorting procedure relied on one feature dimension. On late lists, the words
were always presented in a randomized order (chosen anew for each participant).
In all of the order manipulation conditions, we varied words' font colors and
onscreen locations, as in the feature rich condition.

\paragraph{Constructing feature-sorted lists.}

>>>>

\author{Jeremy R. Manning\textsuperscript{1, *}, Kirsten
Ziman\textsuperscript{1, 2}, Emily Whitaker\textsuperscript{1},\\Paxton C.
Fitzpatrick\textsuperscript{1}, Madeline R. Lee\textsuperscript{1}, 
Allison M Frantz\textsuperscript{1}, \\Bryan J.
Bollinger\textsuperscript{1}, and Andrew C. Heusser\textsuperscript{1,
3}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Princeton
University\\\textsuperscript{3}Akili
Interactive\\\textsuperscript{*}Corresponding author:
jeremy.r.manning@dartmouth.edu}

@@

Each of six \textit{order manipulation} conditions used a different
feature-based sorting procedure to order words on early lists, where each
sorting procedure relied on one relevant feature dimension. All of the
irrelevant features varied freely across words on early lists, in that we did
not consider irrelevant features in ordering the early lists. However, some
features were correlated-- for example, some semantic categories of words
referred to objects that tended to be a particular size, which means that
category and size are not fully independent. On late lists, the words were
always presented in a randomized order (chosen anew for each participant). In
all of the order manipulation conditions, we varied words' font colors and
onscreen locations, as in the feature rich condition.

\paragraph{Defining feature-based distances.} Sorting words according to a
given relevant feature requires first defining a distance function for
quantifying the dissimilarity between each pair of features. This function
varied according to the type of features. Semantic features (category and size)
are \textit{categorical}. For these features, we defined a binary distance
function: two words were considered to ``match'' (i.e., have a distance of 0)
if their labels are the same (i.e., both from the same semantic category or
both of the same size). If two words' labels were different for a given
feature, we defined the words to have a distance of 1 for that feature.
Lexicographic features (length and first letter) are \textit{discrete}. For
these features we defined a discrete distance function. Specifically, we
defined the distance between two words as either the absolute difference
between their lengths, or the absolute distance between their starting letters
in the English alphabet, respectively. For example, two words that started with
the same letter would have a ``first letter'' distance of 0, and words starting
with `J' and `A' would have a first letter distance of 9. Because words'
lengths and letters' positions in the alphabet are always integers, these
discrete distances always take on integer values. Finally, the visual features
(color and location) are \textit{continuous} and \textit{multivariate}, in that
each ``feature'' takes on multiple (positive) real values. We defined the
``color'' and ``location'' distances between two words as the Euclidean
distances between their $(r, g, b)$ color or $(x, y)$ location vectors,
respectively. Therefore the color and location distance measures always take on
positive real values (upper bounded at 441.67 for color, or 27~in for location,
reflecting the distances between the corresponding maximally different
vectors).

\paragraph{Constructing feature-sorted lists.} Given a list of words, a
relevant feature, and each word's value(s) for that feature, we developed a
stochastic algorithm for (noisily) sorting the words. First, we choose a word
uniformly at random from the set of candidates. Next, we compute the distances
between the chosen word's feature(s) and the corresponding feature(s) of all
yet-to-be-presented words. Third, we convert these distances (between the
previously presented word's feature values, $a$, and the candidate word's
feature values, $b$) to similarity scores: \begin{equation}
\mathrm{similarity}(a, b) = \exp\{-\tau \cdot \mathrm{distance}(a, b)\},
\label{eqn:similarity} \end{equation} where $\tau = 1$ in our implementation.
We note that increasing the value of $\tau$ would amplify the influence of
similarity on order, and decreasing the value of $\tau$ would diminish the
influence of similarity on order. Also note that this approach requires $\tau >
0$. Finally, we computed a set of normalized similarity values by dividing the
similarities by their sum: \begin{equation}
\mathrm{similarity}_{\mathrm{normalized}}(a, b) = \frac{\mathrm{similarity}(a,
b)}{\sum_{i=1}^n \mathrm{similarity}(a, i)}, \end{equation} where in the
demoniator, $i$ takes on each of the $n$ feature values of the to-be-presented
words. The resulting set of normalized similarity scores sums to one.

\begin{figure}[tp]
    \centering
        \includegraphics[width=\textwidth]{figures/stick}

\caption{\textbf{Generating stochastic feature-sorted lists.} For a given
feature dimension (e.g., color), we compute the similarity
(Eqn.~\ref{eqn:similarity}) between the feature value(s) of the previous item,
$x$, and all yet-to-be-presented items ($a$ -- $g$). Next, we normalize these
similarity scores so that they sum to one. We lay in sequence a set of
``sticks,'' one for each candidate item, whose lengths are equal to these
normalized similarity scores. Note that the combined lengths of these sticks is
one. To select the next to-be-presented item, we draw a random number, $r$,
from the uniform distibution bounded between 0 and 1 (inclusive). The identity
of the next item is given by the stick adjacent to an indicator that moves
distance $r$ (starting from 0) along the sequence of sticks. In this case, the
next to-be-presented item is $e$. Note that each item's chances of selection is
proportional to its similarity to the previous item, along the given feature
dimension.} \label{fig:stick} 
\end{figure}

As illustrated in Figure~\ref{fig:stick}, we use these normalized similarity
scores to construct a sequence of ``sticks'' that we lay end to end in a line.
Each of the $n$ sticks corresponds to a single to-be-presented word, and the
stick lengths are proportional to the relative similarities between each word's
feature value(s) and the feature value(s) of the just-presented word. We choose
the next to-be-presented word by moving an indicator along the set of sticks,
by a distance chosen uniformly at random on the interval $\left[0, 1\right]$.
We select the word associated with the stick lying next to the indicator to be
presented next. This process continues iteratively (re-computing the similarity
scores and stochastically choosing the next to-be-presented word using the
just-presented word) until all of the words have been presented. The result is
an ordered list that tends to change gradually along the selected feature
dimension.
